\Title

### 2021-03-17 latex/paper/main.tex
Draft 3 of Problem Formulation

In gymnastics terminology, a ``giant" is the motion a gymnast performs to
achieve full rotations around a horizontal bar \cite{usagym_giant}. 
A gymnast will begin by hanging at rest, then swing their legs
appropriately to gain energy over time.
The authors of \cite{pendulum_length_giant_gymnastics} modelled the gymnast as a
variable length pendulum, and studied how the pendulum's length changes as a
function of the orientation of a gymnast's joints.
Labeling the pendulum length by \(r\) and the gymnast's body orientation
by \(\theta\), they discovered experimentally that the value \(\dot{r}/r\) has
the highest impact on the magintude of energy injection; they also discovered
that the peak value of \(\dot{r}/r\) occured at the same value of
\(\dot{\theta}/\theta\) for all gymnasts.
In other words, gymnasts appear to move their legs as a function of their body
angle and velocity; this is the mechanism by which they gain enough energy to
rotate around the bar.

Control theorists typically want a system's actuators to track a fixed
trajectory over time. 
If we were attempting design a gymnastics robot which
could replicate human behaviour, the results of
\cite{pendulum_length_giant_gymnastics} indicate that we should not use this
standard approach.
Instead, we should have our robot's legs follow a function of its
body angle and angular velocity.
That is exactly what we set out to do in this article.

The simplest way to model the gymnast is as a variable-length pendulum, but a
more realistic model is that of a two-link acrobot (Figure \ref{fig:acrobot}).
The acrobot's first link represents the gymnast's torso, the bottom link represents
their legs, and it is actuated at the center joint (the hips).

@@

In this article, our goal is to design a leg controller for the acrobot so that
the resulting motion mimics that of a human performing a giant.
Previous attempts at solving this problem have involved
trajectory tracking, set-point tracking, or energy
pumping regulation (see
\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
).
While these approaches all result in the acrobot performing rotations on the
bar, none of the are particularly human-like.
Wang designed a controller which
forces the hips to track a function of the torso angle using virtual holonomic
constraints \cite{xingbo_thesis};
however, truly human-like hip motion would also take into account the torso velocity.
This style of control action is captured by the method of virtual nonholonomic
constraints.

@@

Here then is the problem we will solve: design a virtual nonholonomic constraint
so that the acrobot performs giant-like motion.
We will rigorously prove the acrobot ``gains energy" according to the following
definition.

@@

Before embarking on this design problem, let us summarize the relevant theory of
VNHCs.

@@

\begin{thm}\label{thm:vnhc-regularity}
    A relation \(h(q,p) = 0\) for system ~\eqref{eqn:simple-hamiltonian}
    is a regular VNHC of order \(k\) if and only if 
    \(dh_{p_a} = \Zmat{k \times k}\) 
    and
    \[
        \rank{\left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right)\simpleB} = k
         ,
    \]
    everywhere on the constraint manifold \(\Gamma\).
\end{thm}
\begin{proof}
    \textbf{TODO: should we make this a ``proof sketch"?}
    Let \(e = h(q,p) \in \R^k\).
    If \(dh_{p_a} \neq \Zmat{k\times k}\) for some \((q,p) \in \Gamma\), 
    then \(\tau\) appears in \(\dot{e}\) and the VNHC is not of relative degree
    \(\{2,\ldots,2\}\). Suppose now that \(dh_{p_a} = \Zmat{k\times k}\).
    Then 
    \(\dot{e} = \mathcal{A}(q,p_u)p - 
     dh_{p_u}\left(1/2~\mathcal{M}(q,p)p + \nabla_{q_u}V(q)\right)\).
    Taking one further derivative provides
    \( \ddot{e} = (\star) - 
        dh_{p_u}\left(1/2~d/dt~\left(\mathcal{M}(q,p)p\right)\right) 
        + \mathcal{A}(q,p_u)[\Zmat{(n-k)\times k};\Id{k}] \tau\),
    where \((\star)\) is a continuous function of \(q\) and \(p\).
    One can further show that
    \(dh_{p_u}\left(1/2~d/dt~\left(\mathcal{M}(q,p)p\right)\right)
        = (\star) + dh_{p_u}\mathcal{M}(q,p)[\Zmat{(n-k)\times k};
        \Id{k}]\tau\).
    Hence,
    \[
       \ddot{e} = (\star) +
       \left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right) \simpleB \tau
        ,
    \]
    which we write as \( \ddot{e} = E(q,p) + H(q,p)\tau\) for appropriate \(E\)
    and \(H\).
    From the definition of regularity, the VNHC \(h\) is regular 
    when \(e\) is of relative degree \(\{2,\ldots,2\}\), which is true 
    if and only if the matrix premultiplying \(\tau\) is nonsingular, and hence
    that \(H\) is invertible. This proves the theorem.
\end{proof}

>>>>

In gymnastics terminology, a ``giant" is the motion a gymnast performs to
achieve full rotations around a horizontal bar \cite{usagym_giant}. 
A gymnast will begin by hanging at rest, then swing their legs
appropriately to gain energy over time.
The authors of \cite{pendulum_length_giant_gymnastics} modelled the gymnast as a
variable length pendulum, and studied how the pendulum's length changes as a
function of the orientation of a gymnast's joints.
Labeling the pendulum length by \(r\) and the gymnast's body orientation
by \(\theta\), they discovered experimentally that the value \(\dot{r}/r\) has
the highest impact on the magintude of energy injection. 
After testing several gymnasts under a variety of experimental conditions, 
they discovered that the peak value of \(\dot{r}/r\) occured at the same fixed
value of \(\dot{\theta}/\theta\) for all gymnasts.
In other words, gymnasts appear to move their legs as a function of their body
angle and velocity when performing giants, not as functions of time. 
Doing so allows them to gain energy and rotate around the bar.

The simplest way to model a gymnast is as a variable-length pendulum, but a
more realistic model is that of a two-link acrobot (Figure \ref{fig:acrobot}).
The acrobot's first link represents the gymnast's torso, the bottom link represents
their legs, and it is actuated at the center joint (the hips).
In this article, our goal is to design a leg controller which provably injects
energy into the acrobot, so that the resulting motion mimics that of a human
performing a giant.

@@

Previous attempts at solving this problem have involved
trajectory tracking, or other feedback methods to stabilize
certain energy levels (see
\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
).
While all these approaches result in the acrobot rotating around the
bar, none of them are particularly human-like.
After all, the results of \cite{pendulum_length_giant_gymnastics} indicate that
a robot's legs should track a function of its body angle and velocity.
In 2016, Wang attempted to design a controller with this in mind \cite{xingbo_thesis}
but was unable to prove the acrobot would gain energy over time.
His approach was a rudimentary version of a recent technique know as the method
of virtual nonholonomic constraint.

@@

Here then is the problem we will solve in this article: design a virtual
nonholonomic constraint so that the acrobot provably gains energy while
performing giant-like motion.
The following is a rigorous definition of ``energy gain".

@@

Before embarking on our design problem, let us summarize the relevant theory of
VNHCs.

@@

\begin{thm}\label{thm:vnhc-regularity}
    A relation \(h(q,p) = 0\) for system ~\eqref{eqn:simple-hamiltonian}
    is a regular VNHC of order \(k\) if and only if 
    \(dh_{p_a} = \Zmat{k \times k}\) 
    and
    \[
        \rank{\left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right)\simpleB} = k
         ,
    \]
    everywhere on the constraint manifold \(\Gamma\).
\end{thm}
\begin{proof}
    Let \(e = h(q,p) \in \R^k\).
    If \(dh_{p_a} \neq \Zmat{k\times k}\) for some \((q,p) \in \Gamma\), 
    then \(\tau\) appears in \(\dot{e}\) and the VNHC is not of relative degree
    \(\{2,\ldots,2\}\). Suppose now that \(dh_{p_a} = \Zmat{k\times k}\).
    Then 
    \(\dot{e} = \mathcal{A}(q,p_u)p - 
     dh_{p_u}\left(1/2~\mathcal{M}(q,p)p + \nabla_{q_u}V(q)\right)\).
    Taking one further derivative provides
    \( \ddot{e} = (\star) - 
        dh_{p_u}\left(1/2~d/dt~\left(\mathcal{M}(q,p)p\right)\right) 
        + \mathcal{A}(q,p_u)[\Zmat{(n-k)\times k};\Id{k}] \tau\),
    where \((\star)\) is a continuous function of \(q\) and \(p\).
    One can further show that
    \(dh_{p_u}\left(1/2~d/dt~\left(\mathcal{M}(q,p)p\right)\right)
        = (\star) + dh_{p_u}\mathcal{M}(q,p)[\Zmat{(n-k)\times k};
        \Id{k}]\tau\).
    Hence,
    \[
       \ddot{e} = (\star) +
       \left(\mathcal{A}(q,p_u) - dh_{p_u}\mathcal{M}(q,p)\right) \simpleB \tau
        ,
    \]
    which we write as \( \ddot{e} = E(q,p) + H(q,p)\tau\) for appropriate \(E\)
    and \(H\).
    From the definition of regularity, the VNHC \(h\) is regular 
    when \(e\) is of relative degree \(\{2,\ldots,2\}\), which is true 
    if and only if the matrix premultiplying \(\tau\) is nonsingular, and hence
    that \(H\) is invertible. This proves the theorem.
\end{proof}
